Mathematics of Reinforcement Learning with Applications to Quantitative Finance

### 2022-08-22 papers/paper_2-col.tex
write up conclusion

%% Abstract should be 200 words or less
\begin{abstract}
\noindent This paper provides a new way of pricing in the car insurance
industry, which is, using reinforcement learning algorithms to train pricing 
agents to promote proper price, such that our customer is most likely to 
purchase our product, and we also get the highest profit and a better 
portfolio. 
We first introduce general reinforcement learning knowledge, including how 
we model the interaction between agent and environment using MDP, and the 
mathematics behind those methods to solve the reinforcement learning problem, 
especially the Actor-Critic algorithm.
Then we start building the environment, and use ANNs as our pricing models, 
and writing algorithms that are specific to our financial problem. 
In the end, we will compare and discuss different algorithms in terms of their
rate of convergence to a stable reward.
\end{abstract}

@@

\section{Actor-Critic, Reinforce, and PPO}
\subsection{Algorithms}
\noindent To further show how Actor-Critic algorithm performs, we compare it 
to another two popular algorithms: Reinforce 
(Algorithm~\ref{algorithm:reinforce}) and PPO (Algorithm~\ref{algorithm:ppo}).
\begin{algorithm}
    \caption{Reinforce for financial problem}\label{algorithm:reinforce}
    \begin{algorithmic}[1]
        \State Initialize ANN policy parameterization
        \(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
        \State Step sizes \(\alpha_{\bm{\theta}}>0\)
        \Loop \hspace{0.3mm} (for each episode)
        \State Reset the environment
        \State Generate an episode 
        \(S_0, C_0, R_1, ..., S_{T-1}, C_{T-1}, R_T\),
        following \(\pi\left(\cdot\middle|\cdot, \bm{\theta}\right)\)
        \For{\(t=0, 1, ..., T\)}
        \State \(G \gets \sum_{k=t+1}^{T} \gamma^{k-t-1}R_k\)
        \State \(\bm{\theta} \gets \bm{\theta} + \alpha_{\bm{\theta}}\gamma^t
        G\nabla{\ln{\pi\left(C_t\middle|S_t, \bm{\theta}\right)}}\)
        \EndFor
        \EndLoop
    \end{algorithmic}
\end{algorithm}

@@

\subsection{Training result}
\noindent To make better comparison, the policy function of the two algorithms
is implemented the same way as we did in Actor-Critic algorithm, and the same
is true for the value function used in PPO method.
\vspace{3mm}\newline Figure~\ref{fig:3-algorithms-out} shows the training 
result of these three algorithms (300 moving average reward vs iteration), 
with 600 iterations (5 episodes per iteration), \(T=300\), \(\gamma=0.99\),
\(\alpha_\theta=\num{3e-4}\), \(\alpha_{\bm{w}}=\num{3e-10}\) (if applicable),
\(\epsilon=0.5\) (if applicable).
\begin{figure*}
    \centering\includegraphics[scale=0.75]{out}
    \caption{Training Result Comparison for Three Algorithm}
    \label{fig:3-algorithms-out}
\end{figure*}
As is shown in Figure~\ref{fig:3-algorithms-out}, the Actor-Critic algorithm 
converges the fastest, within 50 iterations, followed by Reinforce algorithm, 
which converges within about 250 iterations, despite a small drop of moving 
average around iteration 300, which is quickly corrected by the algorithm 
within 50 iterations. PPO algorithm converges the slowest, with about 300 
iterations. Despite of the speed of convergence, the moving average rewards 
of convergence for these three algorithms are nearly the same (around 400), 
this is as expected.

\section{Conclusion}
\noindent Provide a brief conclusion (3-4 lines).

>>>>

%% Abstract should be 200 words or less
\begin{abstract}
\noindent This paper provides a new way of pricing in the car insurance
industry, which is, using reinforcement learning algorithms to train pricing 
agents to promote proper price, such that our customer is most likely to 
purchase our product, and we also get the highest profit and a better 
portfolio. 
We first introduce general reinforcement learning knowledge, including how 
we model the interaction between agent and environment using Markov decision
process, and the mathematics behind those methods to solve the reinforcement 
learning problem, especially the Actor-Critic algorithm.
Then we start building the environment, and use artificial neural networks as 
our pricing models, and writing algorithms that are specific to our financial 
problem. 
In the end, we will compare the Actor-Critic algorithms with other popular 
algorithms including REINFORCE and PPO in terms of their rate of convergence 
to a stable reward.
\end{abstract}

@@

\section{Actor-Critic, Reinforce, and PPO}
\subsection{Algorithms}
\noindent To further show how Actor-Critic algorithm performs, we compare it 
to another two popular algorithms: Reinforce 
(Algorithm~\ref{algorithm:reinforce}) and PPO (Algorithm~\ref{algorithm:ppo}).
The policy function of the two algorithms is implemented the same way as we 
did in Actor-Critic algorithm, and the same is true for the value function 
used in PPO method.
\begin{algorithm}
    \caption{Reinforce for financial problem}\label{algorithm:reinforce}
    \begin{algorithmic}[1]
        \State Initialize ANN policy parameterization
        \(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
        \State Step sizes \(\alpha_{\bm{\theta}}>0\)
        \Loop \hspace{0.3mm} (for each episode)
        \State Reset the environment
        \State Generate an episode 
        \(S_0, C_0, R_1, ..., S_{T-1}, C_{T-1}, R_T\),
        following \(\pi\left(\cdot\middle|\cdot, \bm{\theta}\right)\)
        \For{\(t=0, 1, ..., T\)}
        \State \(G \gets \sum_{k=t+1}^{T} \gamma^{k-t-1}R_k\)
        \State \(\bm{\theta} \gets \bm{\theta} + \alpha_{\bm{\theta}}\gamma^t
        G\nabla{\ln{\pi\left(C_t\middle|S_t, \bm{\theta}\right)}}\)
        \EndFor
        \EndLoop
    \end{algorithmic}
\end{algorithm}

@@

\subsection{Training result}
\noindent Figure~\ref{fig:3-algorithms-out} shows the training 
result of these three algorithms (300 moving average reward vs iteration), 
with 600 iterations (5 episodes per iteration), \(T=300\), \(\gamma=0.99\),
\(\alpha_\theta=\num{3e-4}\), \(\alpha_{\bm{w}}=\num{3e-10}\) (if applicable),
\(\epsilon=0.5\) (if applicable).
\begin{figure*}
    \centering\includegraphics[scale=0.75]{out}
    \caption{Training Result Comparison for Three Algorithm}
    \label{fig:3-algorithms-out}
\end{figure*}
As we can see, the Actor-Critic algorithm converges the fastest, within 50 
iterations, followed by REINFORCE algorithm, which converges within 250 
iterations, despite a small drop of moving average around iteration 300, 
which is quickly corrected by the algorithm within 50 iterations. PPO 
algorithm converges the slowest, with about 300 iterations.

\section{Conclusion}
\noindent To summarize, the moving average rewards after convergence of these 
three algorithms are nearly the same (around 400), which is as expected. 
The Actor-Critic algorithm performs really well compared to anther two 
algorithms, it has highest converge rate, the reward is also very stable after 
convergence. 
In the mean time, the slow convergence for PPO algorithm is expected since it 
uses a clamp function to prevent model from changing too fast in a single 
update, although at current stage it seems to be unnecessary for our specific 
problem setup. 
Further investigation needs to be done for the unexpected small drop of the 
reward for the REINFORCE algorithm.
